---
title: "Iris Flower Data Analysis"
author: Pooja Kalia
date: "2026-09-17"
output: html_document
---
# Purpose

This document explains the R script used to analyse the built-in iris flower dataset. The analysis summarizes flower measurements, compares average sepal length among species and creates a scatter plot.

# Inputs

The project explains the R script used to analyse the built-in iris flower dataset. It contains 150 observations and the following variables:

- Sepal Length
- Sepal Width
- Petal Length
- Petal Width
- Species

No external files or additional packages are required.

# Load the dataset

```{r load-data}
data(iris)
head(iris)
```

# Dataset information

```{r dataset-information}
str(iris)
summary(iris)
```

# Average Sepal Length

The following code calculates the average sepal length for each species.

```{r average-Sepal-Length}
aggregate(Sepal.Length ~ Species, data = iris, FUN = mean)
```

# Scatter plot

```{r scatter-plot}
plot(
iris$Sepal.Length,
iris$Petal.Length,
main = "Sepal Length vs Petal Length",
xlab = "Sepal Length",
ylab = "Petal Length",
col = iris$Species,
pch = 19
)
legend(
   "topleft",
   legend = levels(iris$Species),
   col = 1:3,
   pch = 19
)
```
   
# Output
   
   The script produces:
   
   - A preview of the iris Dataset
   - Dataset structure and summary statistics
   - Average sepal length for each species
   - A scatter plot comparing sepal length and petal length
   
# Conclusion
   
Virginica has the highest average sepal length, followed by versicolor and setosa. The scatter plot indicates a positive relationship between sepal length and petal length.